---
title: "Econ 424 Project Group 7"
output:
  pdf_document: default
  html_notebook: default
  html_document:
  df_print: paged
---

```{r, echo = FALSE}
options(digits=3, width=70)
# load packages
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(IntroCompFinR))
suppressPackageStartupMessages(library(PerformanceAnalytics))
suppressPackageStartupMessages(library(quantmod))
suppressPackageStartupMessages(library(boot))
suppressPackageStartupMessages(library(corrplot))
suppressPackageStartupMessages(library(xtable))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gridExtra))

resetPar = par()
```


# Data

```{r, echo=FALSE}
# retrieve data for questions
last.date = as.Date("2023-12-31")
# last.date = Sys.Date()
first.date = last.date - years(10)
project.symbols = c("EWS","EWY","EWT","EWH")
project.prices <- 
  getSymbols(project.symbols, src = 'yahoo', 
             from = as.character(first.date), 
             to = as.character(last.date),
             auto.assign = TRUE, 
             warnings = FALSE) %>% 
  map(~Ad(get(.))) %>% 
  reduce(merge) %>%
  `colnames<-`(project.symbols)
projectPrices = to.monthly(project.prices, OHLC=FALSE)  # Group 7 you should continue to use daily data and skip this
```

```{r, echo=FALSE}
# calculate simple returns
projectReturns = na.omit(Return.calculate(projectPrices, method = "discrete"))
ret.mat = coredata(projectReturns)
```

## Historical performance

```{r, echo=FALSE}
chart.CumReturns(projectReturns, main="Example data",
                 wealth.index = TRUE, legend.loc = "topleft")
```

* Starbucks performed best over the full sample but Microsoft was best before the dot-com bust

## Estimated inputs to portfolio theory: full sample

```{r}
nobs = nrow(projectReturns)
muhat.vals = colMeans(projectReturns)
sigmahat.vals = apply(projectReturns,2,sd)
cov.mat = var(projectReturns)
cor.mat = cor(projectReturns)
```

Full sample mean values

```{r}
muhat.vals
```

Full sample volatility values

```{r}
sigmahat.vals
```

Full sample correlation matrix

```{r}
cor.mat
```

## Risk-return tradeoff: full sample

```{r}
plot(sigmahat.vals, muhat.vals,  
     ylim=c(0, 0.02), xlim=c(0, 0.1), 
     ylab=expression(mu[p]), xlab=expression(sigma[p]), 
     pch=16, col="blue", cex=1, cex.lab=1)     
text(sigmahat.vals, muhat.vals, labels=names(muhat.vals), pos=1, cex = 1.1)
```

* Starbucks is best, followed by Microsoft and EWYstrom

## Risk-return tradeoff: by year

```{r, echo=FALSE}
muhatByYear = rollapply(projectReturns, width=12, by=12, FUN=mean, align="right")
muhatByYear = na.omit(muhatByYear)

sigmahatByYear = rollapply(projectReturns, width=12, by=12, FUN=sd, align="right")
sigmahatByYear = na.omit(sigmahatByYear)

plot(sigmahat.vals, muhat.vals,  
     ylim=c(-0.02, 0.06), xlim=c(0, 0.20), 
     ylab=expression(mu[p]), xlab=expression(sigma[p]), 
     pch=16, col="blue", cex=1.1, cex.lab=1)     
text(sigmahat.vals, muhat.vals, labels=names(muhat.vals), pos=1, cex = 1.1)

for (i in 1:nrow(muhatByYear)) {
  points(coredata(sigmahatByYear[i, ]), coredata(muhatByYear[i, ]), pch=16, col=i)
  text(coredata(sigmahatByYear[i, ]), coredata(muhatByYear[i, ]), labels=names(muhat.vals), 
         pos=4, cex = 0.5)
}

```

* Big blue circles are full sample estimates
* Small colored circles are values for a particular year
* Risk return tradeoff changes drastically over time!

## 24-month rolling means and volatilities

Use **zoo** function `rollapply()` to compute 24-month rolling means and volatilities

```{r}
roll.muhat.EWS = rollapply(projectReturns[,"EWS"], width=24,
                       FUN=mean, align="right")
roll.sigmahat.EWS = rollapply(projectReturns[,"EWS"],width=24,
                          FUN=sd, align="right")
roll.muhat.EWY = rollapply(projectReturns[,"EWY"], width=24,
                            FUN=mean, align="right")
roll.sigmahat.EWY = rollapply(projectReturns[,"EWY"],width=24,
                               FUN=sd, align="right")
roll.muhat.EWT = rollapply(projectReturns[,"EWT"], width=24,
                            FUN=mean, align="right")
roll.sigmahat.EWT = rollapply(projectReturns[,"EWT"],width=24,
                               FUN=sd, align="right")
roll.muhat.EWH = rollapply(projectReturns[,"EWH"], width=24,
                            FUN=mean, align="right")
roll.sigmahat.EWH = rollapply(projectReturns[,"EWH"],width=24,
                               FUN=sd, align="right")
```

## 24-month rolling means and volatilities

```{r, echo=FALSE}
par(mfrow=c(2,2))
# EWS
  plot.zoo(merge(roll.muhat.EWS, roll.sigmahat.EWS, projectReturns[,"EWS"]), 
          plot.type="single",
          main="EWS",ylab="returns",
          lwd=2, col=c("blue","orange","black"))
  abline(h=0)

 
# EWY
  plot.zoo(merge(roll.muhat.EWY, roll.sigmahat.EWY, projectReturns[,"EWY"]), 
          plot.type="single",
          main="EWY",ylab="returns",
          lwd=2, col=c("blue","orange","black"))
  abline(h=0)  
  
# EWT
  plot.zoo(merge(roll.muhat.EWT,roll.sigmahat.EWT, projectReturns[,"EWT"]), 
          plot.type="single",
          main="EWT",ylab="returns",
          lwd=2, col=c("blue","orange","black"))
  abline(h=0)
  
# EWH
  plot.zoo(merge(roll.muhat.EWH,roll.sigmahat.EWH, projectReturns[,"EWH"]), plot.type="single",
       main="EWH",ylab="returns",
       lwd=2, col=c("blue","orange","black"))
  abline(h=0)
  
# blank
  plot.zoo(roll.muhat.EWS, type="n", ylab="")
  legend(x="center",legend=c("Rolling mean","Rolling sd", "Monthly returns"),
         lwd=2, col=c("blue","orange","black"))
par(mfrow=c(1,1))

```

## 24-month rolling correlations

Here there are three pair-wise correlations: EWS-EWY, EWS-EWT, and EWY-EWT. The following function extracts these pair-wise correlations from the estimated correlation matrix.

```{r}
roll.cor = function(x) {
  cor.hat = cor(x, use = "pairwise.complete.obs")  
  cor.vals = cor.hat[lower.tri(cor.hat, diag = FALSE)]
  
  # Generate proper names for each pair
  tickers <- colnames(x)
  pair_names <- combn(tickers, 2, FUN = function(pair) paste(pair, collapse = "."))
  names(cor.vals) <- pair_names
  
  return(cor.vals)
}

```

Compute all pair-wise 24-month rolling correlations using `rollapply()`

```{r}
roll.cor.vals = rollapply(projectReturns, width=24,
                          by.column=FALSE,
                          FUN=roll.cor, 
                          align="right")
```



## 24-month rolling correlations

```{r}
my.panel <- function(...) {
  lines(...)
  abline(h=0)
}

colnames(roll.cor.vals) <- colnames(roll.cor.vals, do.NULL = FALSE, prefix = "Pair")

# Re-run the plotting function
plot.zoo(roll.cor.vals, panel=my.panel, main="Rolling Pairwise Correlations",
     lwd=2, col="blue", ylim=c(-0.25,1))

```

* Rolling correlations change over time but not too much

## Global minimum variance portfolio: full sample

```{r}
gmin.full = globalMin.portfolio(er=muhat.vals,cov.mat=cov.mat)
gmin.full
```

* Full sample global minimum variance portfolio is roughly equally weighted

## 24-month rolling global minimum variance portfolio

Now we compute the global minimum variance portfolio over 24-month rolling window. First, we need to write a function to return the global minimum variance portfolio weights, means and volatilities for each rolling window. We use the **IntroCompfinR** function `globalMin.portfolio()` to do the calculations.

```{r}
# function to compute rolling global minimum variance portfolios
rollGmin = function(x) {
	mu.hat = colMeans(x)
	cov.hat = var(x)
	gmin = globalMin.portfolio(er=mu.hat,cov.mat=cov.hat)
	ans = c(gmin$er,gmin$sd,gmin$weights)
	names(ans)[1:2] = c("er","sd")
	return(ans)
}
```

Now use `rollapply()` with `rollGmin()` to do the calculations.

```{r}
# use rollapply() to compute rolling portfolios
roll.gmin = rollapply(projectReturns, width=24,
                      by.column=FALSE,align="right",
                      FUN=rollGmin)

# show first three non NA values
na.omit(roll.gmin)[1:3, ]
```


## 24-month rolling global minimum variance portfolio weights

```{r, echo=FALSE}
plot.zoo(roll.gmin[,3:6], main="",
         plot.type="single", col=1:4, lwd=3, ylab="weight")    
abline(h=0)
legend(x="bottomleft", legend=colnames(roll.gmin[,3:6]),
       lty=rep(1,4), col=1:4, lwd=3)
```

* Rolling portfolio weights change a lot over time
* Look at how the weights in Microsoft and Starbucks change over time

## 24-month rolling global minimum variance portfolio weights

```{r, echo=FALSE}
chart.StackedBar(na.omit(roll.gmin[,3:6]))
```

* Stacked bar chart created using the PerformanceAnalytics function `chart.StackedBar()` shows changing weights more clearly

## 24-month rolling global minimum variance portfolio means and volatilities

```{r, echo=FALSE}
plot.zoo(roll.gmin[,1:2],plot.type="single",ylab="percent",
         main="",
         col=c("black","blue"),lwd=3)
abline(h=0)
legend(x="topleft",legend=c("Rolling mean","Rolling sd"),
       lty=rep(1,2),col=c("black","blue"),lwd=3)
```

* Notice how mean and volatility move in opposite directions: when mean goes down vol goes up!

## Efficient portfolio with target return $\mu_p = 0.03$ : Full Sample

```{r}
eport.03 = efficient.portfolio(er=muhat.vals,cov.mat=cov.mat,
                                target.return=0.03)
eport.03

```

Full sample portfolio is a long-short portfol with long positions in Microsoft and Starbucks.

## 24-month rolling efficient portfolios with target return 3%

Now we compute the efficient portfolio with target return 3% over 24-month rolling window. First, we need to write a function to return the efficient portfolio weights, means and volatilities for each rolling window. We use the **IntroCompfinR** function `efficient.portfolio()` to do the calculations.

```{r}
# function to compute rolling efficient portfolios
rollefficient = function(x,target=0.03) {
	mu.hat = colMeans(x)
	cov.hat = var(x)
	eport = efficient.portfolio(er=mu.hat,
                              cov.mat=cov.hat,
	                            target.return=target)
	ans = c(eport$er,eport$sd,eport$weights)
	names(ans)[1:2] = c("er","sd")
	return(ans)
}
```

Now use `rollapply()` with `rollefficient()` to do the calculations.

```{r}
# use rollapply() to compute rolling portfolios
roll.eport = rollapply(projectReturns, width=24,
                       by.column=F,align="right",
                       FUN=rollefficient)
```

## 24-month rolling efficient portfolios

```{r, echo=FALSE}
plot.zoo(roll.eport[,3:6],main="",
     plot.type="single", ylab="weight", col=1:4,lwd=3)
abline(h=0)
legend(x="bottomleft",legend=colnames(projectReturns),
       lty=rep(1,4),col=1:4,lwd=3)
```
## 24-month rolling efficient portfolios

```{r, echo=FALSE}
chart.StackedBar(na.omit(roll.eport[,3:6]))
```
## 24-month rolling global minimum variance portfolio means and volatilities

```{r, echo=FALSE}
plot.zoo(roll.eport[,1:2], plot.type="single", ylab="percent",
         main="", ylim=c(0, 0.35),
         col=c("black","blue"),lwd=3)
abline(h=0)
legend(x="topleft",legend=c("Target er=0.03","Rolling sd"),
       lty=rep(1,2),col=c("black","blue"),lwd=3)

```